---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# coefbounds

Nonparametric Bounds for Regression with Interval-Censored Outcomes

Author: Brenton Kenkel, Vanderbilt University


## Installation

**coefbounds** is not yet available on CRAN.  To install directly from GitHub:

```r
if (!require("devtools")) {
    install.packages("devtools")
    library("devtools")
}

devtools::install_github("brentonk/coefbounds", build_vignettes = TRUE)
```


## Note

**coefbounds** is in an early stage of development.  Backward-incompatible API changes are possible until version 1.0.0 is released.


## Usage

Below are some simple usage examples.  For more fleshed-out examples, see the package vignette: `vignette("introduction", package = "coefbounds")`.

```{r load-package, message = FALSE}
library("coefbounds")
```

I'll begin by generating some simple data with an interval-censored outcome.

```{r sim-data}
set.seed(97)
x1 <- rnorm(100)
x2 <- rnorm(100)
y <- 1 - x1 + x2 + rnorm(100)
lwr <- floor(y)
upr <- ceiling(y)
```

To estimate coefficient bounds, use `lm_bounds()` with a formula of the form `yl + yu ~ x1 + x2 + ...`, where `yl` is the lower bound on each response value and `yu` is the upper bound.

```{r full-bounds}
fit_full <- lm_bounds(lwr + upr ~ x1 + x2, boot = 100)
fit_full
```

For inference, make sure `lm_bounds()` is run with `boot > 0` and use `interval_hypothesis()` or `confint()`.

```{r int-hyp}
interval_hypothesis(fit = fit_full,
                    term = "x1",
                    interval = c(0, 0),
                    type = "subset")
interval_hypothesis(fit = fit_full,
                    term = "x2",
                    interval = c(0.7, 1.6),
                    type = "equal")
```

```{r confint}
confint(fit_full, level = 0.99)
```


## To Do

* Allow clustering in critical value bootstrap.
* Enable parallel processing in bootstrap.
